Simulating dynein's powerstroke using Brownian dynamics

### 2017-05-11 thesis/latex/thesis-old.tex
Thesis: Move over to a template. thesis-old.tex has the old thesis in it.

\blockquote{Dynein is a motor protein which transports cargo along tracks inside the cell. Like related motor proteins kinesin and myosin, dynein uses cellular energy to take steps with its two foot domains. Unlike the more regularly-stepping kinesin or myosin, dynein's stepping pattern is very random: it can take steps between zero and 60nm in both the forwards and backwards directions. It is believed that dynein takes such broad, stochastic steps because of its large size and several elastic regions, which make it more influenced by Brownian motion. To test this, we model the motor as a 2D system of springy hinges, then simulate this model using Brownian dynamics. Preliminary results indicate such a model is capable of taking steps between zero and 25nm. These results give hope that, with further tweaking, the model may be able to generate both larger and backwards steps.}

@@

Cells need to create asymmetric, heterogeneous structures inside themselves in order to function. However, as per the above quote, diffusive processes cannot create this asymmetry. How then do cells create heterogeneous structures they need in order to exist? Motor proteins are part of the answer. Motor proteins are structures which use energy to create directed motion in cells. In broad terms, they are mechanical creators of asymmetry. This thesis will explore a hypothesis on how one such motor protein, Dynein, generates directed motion.\\

@@

\subsection{Motivation}
Motor proteins are cellular structures which use chemical energy to do work and generate motion. Kinesin, myosin and dynein are the three main motor proteins in cells. Though these proteins have significantly different ancestors \ref{cite}, they all evolved into a single similar structure. In biology when a structure is evolved three separate times, it is a sign that nature has found an efficient way for solving a problem. How, then, do motor proteins work?\\

The basic form is simple. Kinesin, myosin and dynein each use two binding domains to attach to a surface. These binding domains are connected via a linker domain. Each motor protein is an engine which undergoes cycles, where each cycle moves the motor across a surface. In a single cycle a binding domain will unbind from a surface, tense its foot to kick forward, rebind, then untense its foot to pull the whole motor forward. This pattern is repeated by both feet to generate motion \cite{kinesin-cycle} \cite{myosin-cycle}.\\

Although the three motor protein families operate using the same stepping cycle, their behavior differs significantly. Kinesin and myosin are predictable motors which step in the forwards direction with a regular step size [get kinesin, myosin stepping papers]. In contrast, dynein's stepping pattern varies wildly in both size and direction. The motor steps forward on average, but takes steps backwards and to the side quite regularly \cite{weihongpaper}. How can motors which implement the same release, lever, bind, lurch stepping mechanism behave so differently?\\

One possibility is that dynein's size and organization are responsible for the motor's odd stepping pattern. Dynein is much larger than kinesin or myosin, meaning its feet may need to move significantly further over each cycle. Due to the random nature of Brownian motion from water particle collisions, there may be significantly more variance in the path taken by dynein's feet compared to other motor feet. This could explain the large variance in step size. Another contributing factor could be that dynein is so large that it can twist and contort during its step, allowing varying and even backwards steps, which a small motor would be too tense to achieve.\\

This thesis describes an experiment which aims to test whether Brownian motion and flexibility are what give rise to dynein's unique stepping pattern. To do this, a mathematical model of a dynein-like motor is constructed which exhibits flexibility and feels Brownian forces. This model is then simulated, and its behavior is compared to the behavior of real dynein.\\

>>>>

\blockquote{Dynein is a motor protein which transports cargo along tracks inside the cell. Like related motor proteins kinesin and myosin, dynein uses cellular energy to take steps with its two foot domains. Unlike kinesin or myosin, dynein's stepping pattern is highly varied: it can take steps between zero and 60nm in both the forwards and backwards directions. It is believed that dynein takes such broad, stochastic steps because its large size and several elastic regions make it more influenced by Brownian motion. To test this, we model the motor as a 2D system of elastic hinges, then simulate this model using Brownian dynamics. Preliminary results indicate such a model is capable of taking steps between zero and 25nm. These results give hope that, with further tweaking, the model may be able to generate both larger steps and backwards steps.}

@@

Cells need to create asymmetric, heterogeneous structures inside themselves in order to function. However, diffusive processes cannot create this asymmetry. How then do cells create the heterogeneous structures they need to exist? Motor proteins are part of the answer. Motor proteins are structures which use energy to create directed motion in cells. In broad terms, they are mechanical creators of asymmetry. This thesis explores a hypothesis about how one such motor protein, dynein, generates directed motion.\\

@@

\subsection{Motivation}
Motor proteins are cellular structures which use chemical energy to do work and generate motion. Kinesin, myosin and dynein are the three main motor proteins in cells. Though these proteins have significantly different ancestors \ref{cite}, they all evolved into a single similar structure. In biology when a structure is evolved three separate times, it is a sign that nature has found an efficient way of solving a problem. How, then, do motor proteins work?\\

The basic process is simple. Kinesin, myosin and dynein each use two binding domains to attach to a surface. These binding domains are connected via a linker domain. Each motor protein is an engine which undergoes cycles, where each cycle moves the motor across a surface. In a single cycle a binding domain unbinds from a surface, tenses its foot to kick forward, rebinds, then untenses its foot to pull the whole motor forward. This pattern is repeated by both feet to generate motion \cite{kinesin-cycle} \cite{myosin-cycle}.\\

Although the three motor protein families operate using the same stepping cycle, their behaviors differ significantly. Kinesin and myosin are predictable motors which step in the forward direction with a regular step size [get kinesin, myosin stepping papers]. In contrast, dynein's stepping pattern varies in both size and direction. The motor can take 64nm steps forward or 32nm steps backward, and often steps sideways as well \cite{weihongpaper}. How can motors which implement the same release, lever, bind, lurch stepping mechanism behave so differently?\\

One possibility is that dynein's size and organization are responsible for the motor's odd stepping pattern. Dynein is much larger than kinesin or myosin, so its feet may need to move significantly further during each cycle. Due to the random nature of Brownian motion from water particle collisions, there may be significantly more variance in the path taken by dynein's feet compared to other motor feet. This could explain the large variance in step size. Another contributing factor could be that dynein is so large that it can twist and contort during its step, allowing varying and even backwards steps, which a small motor would be too stiff to achieve.\\

This thesis describes an experiment which aims to test whether Brownian motion and flexibility cause dynein's unique stepping pattern. To do this, we construct a mathematical model of the dynein motor which exhibits flexibility and feels Brownian forces. We then simulate the model and compare its behavior to real dynein.\\
